\textbf{Efficient Implementation Strategies for Block Ciphers on ARMv8}\\ {\footnotesize Bachelorarbeit}

### 2023-04-10 Paper/chapters/implementation.tex
camellia introduction

These kinds of optimizations are pervasive when programming using higher level
languages like C and modern-day compilers more often than not outperform even
carefully handcrafted assembly.

>>>>

These kinds of optimizations are pervasive when programming using higher level
languages like C and modern-day compilers more often than not outperform
handcrafted assembly.

### 2023-04-10 Paper/chapters/introduction.tex
camellia introduction

\texttt{GIFT} comes in two variants; \verb|GIFT-64| working with 64-bit blocks
and \texttt{GIFT-128} working with 128-bit blocks. In both cases, the key is 128
bits long. The design is a very simple, round-based substitution-permutation
network (SPN). One round consists in a sequential application of the confusion
layer by means of 4-bit S-boxes and subsequent diffusion through bit
permutation. After the bit permutation, a round key is added to the cipher
state and the single round is complete. \texttt{GIFT-64} uses 28 rounds while
\texttt{GIFT-128} uses 40 rounds.

@@

The input of \texttt{GIFT} is split into 4-bit nibbles which are then fed into
16 S-boxes for \texttt{GIFT-64} and 32 S-boxes for \texttt{GIFT-128}. The S-box
$S:\F{4}\rightarrow \F{4}$ is defined as follows:

@@

The permutation $P$ works on individual bits and maps bit $b_i$ to $b_{P(i)},
i\in\{0,1,\dots,n-1\}$. The different permutations for \texttt{GIFT-64} and
\texttt{GIFT-128} can be expressed by:

@@

The last step of each round consists in XORing a round key $R_i$ to the cipher
state. The new cipher state $s_{i+1}$ after each full round is therefore given
by

\[
    s_{i+1}=P(S(s_i))\oplus R_i
\]

@@

Round key extraction differs for \texttt{GIFT-64} and \texttt{GIFT-128}. Let
$K=k_7||k_6||\dots||k_0$ denote the $128$-bit key state.

\paragraph{GIFT-64}. We extract two 16-bit words $U||V=k_1||k_0$ from the key
state. These are then added to the 64-bit round key: $R_{4i+1}\leftarrow
u_i,R_{4i}\leftarrow v_i$.

\paragraph{GIFT-128}. We extract two 32-bit words $U||V=k_5||k_4||k_1||k_0$ from
the key state. These are then added to the 128-bit round key: $R_{4i+2}\leftarrow
u_i,R_{4i+1}\leftarrow v_i$.

In both cases, we additionally XOR a round constant $C=c_5c_4c_3c_2c_1c_0$ to
bit positions $n-1,23,19,15,11,7,3$. The round constants are generated using a
6-bit affine linear-feedback shift register and have the following values:\\

@@

The key state is then updated by setting $k_1\leftarrow k_1\ggg 2$,
$k_0\leftarrow k_0\ggg 12$ and rotating the new state $32$ bits to the right:

@@

>>>>

\texttt{GIFT} comes in two variants; \verb|GIFT-64| working with 64-bit blocks
and GIFT-128 working with 128-bit blocks. In both cases, the key is 128
bits long. The design is a very simple, round-based substitution-permutation
network (SPN). One round consists in a sequential application of the confusion
layer by means of 4-bit S-boxes and subsequent diffusion through bit
permutation. After the bit permutation, a round key is added to the cipher
state and the single round is complete. GIFT-64 uses 28 rounds while
GIFT-128 uses 40 rounds.

@@

The input of \texttt{GIFT} is split into 4-bit nibbles which are then fed into
16 S-boxes for GIFT-64 and 32 S-boxes for GIFT-128. The S-box
$S:\F{4}\rightarrow \F{4}$ is defined as follows:

@@

The permutation $P$ works on individual bits and maps bit $b_i$ to $b_{P(i)}$.
The different permutations for GIFT-64 and GIFT-128 can be expressed by:

@@

The last step of each round consists in XORing a round key $R_{i}$ to the cipher
state. The new cipher state $x_{i+1}$ after each full round is therefore given
by

\[
    x_{i+1}=P(S(x_i))\oplus R_i
\]

@@

Round key extraction differs for GIFT-64 and GIFT-128. Let
$K_{(128)}=k_7||k_6||\dots||k_0$ denote the 128-bit key state.

\paragraph{GIFT-64}. We extract two words $U_{(16)}||V_{(16)}=k_1||k_0$ from the key
state. These are then added to round key $R_{(64)}$: $R_{4i+1}\leftarrow
U_i,R_{4i}\leftarrow V_i$.

\paragraph{GIFT-128}. We extract two words
$U_{(32)}||V_{(32)}=k_5||k_4||k_1||k_0$ from the key state. These are then
added to round key $R_{(128)}$: $R_{4i+2}\leftarrow U_i,R_{4i+1}\leftarrow
V_i$.

In both cases, we additionally XOR a round constant $C_{(6)}$ to bit positions
$n-1,23,19,15,11,7,3$. The round constants are generated using a 6-bit affine
linear-feedback shift register and have the following values:\\

@@

The key state is then updated by individually rotating $k_1$ and $k_0$ and
rotating the new state $32$ bits to the right:

@@

Camellia\cite{camellia:2001} is a block cipher jointly developed by NTT and
Mitsubishi Electric Corporation and first published in 2001. Following AES
specifications, it is able to encrypt 128-bit blocks using either 128-, 196- or
256-bit keys and claims to possess similar performance and security levels as
the AES finalists.

\subsubsection{Encryption overview}

The encryption process has an 18-round Feistel structure for 128-bit keys and a
24-round Feistel structure for 192/256-bit key and employs key whitening to
increase security. First, subkeys $kw_{t(64)}(t=0,1,2,3)$,
$k_{u(64)}(u=1,2,\dots,(17|23)$ and $kl_{v(64)}(v=0,1,2,3)$ are generated from
the master key. Then, pre-whitening keys are applied to the left and right half
of the plaintext $m_{(128)}=L_{(64)}||R_{(64)}$:

\[
    L\leftarrow L\oplus kw_0, R\leftarrow R\oplus kw_1
\]

The next steps differ for 128-bit and 192/256-bit keys:

\begin{table}[h!]
    \centering
    \caption{Camellia encryption}
    \begin{tabular}{lll}
        \toprule
        Round $r$ & 128-bit keys & 192/256-bit keys \\
        \midrule
        0-4 & $FE(L,R,k_r)$ & $FE(L,R,k_r)$ \\
        \midrule
        5 & {$\!\begin{aligned}&FE(L,R)\\&FLA(L,R,kl_0,kl_1)\end{aligned}$} & {$\!\begin{aligned}&FE(L,R)\\&FLA(L,R,kl_0,kl_1)\end{aligned}$} \\
        \midrule
        6-10 & $FE(L,R,k_r)$ & $FE(L,R,k_r)$ \\
        \midrule
        11 & {$\!\begin{aligned}&FE(L,R)\\&FLA(L,R,kl_2,kl_3)\end{aligned}$} & {$\!\begin{aligned}&FE(L,R)\\&FLA(L,R,kl_2,kl_3)\end{aligned}$} \\
        \midrule
        12-16 & $FE(L,R,k_r)$ & $FE(L,R,k_r)$ \\
        \midrule
        17 & & {$\!\begin{aligned}&FE(L,R)\\&FLA(L,R,kl_4,kl_5)\end{aligned}$} \\
        \midrule
        18-23 & & $FE(L,R,k_r)$ \\
        \bottomrule
    \end{tabular}
\end{table}

Finally, $R$ and $L$ are concatenated and XORed with the post-whitening keys to
obtain the cipher text $c_{(128)}$:

\[
    c=(R\oplus kw_2)||(L\oplus kw_3)
\]

\subsubsection{Components}

$FE$ and $FLA$ are macros for applying one round of the Feistel structure and
$FL$ functions and are defined as follows:

### 2023-04-10 Paper/paper.tex
remove implementation.tex, fancyhdr
begin work on benchmark
benchmark
bibliography no ul
implementation chapter
camellia introduction

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{ulem} % underline
\usepackage{float} % H float specifier
\usepackage{hyperref} % hyperlinks
\usepackage{tabularx} % tabular with column specifier X to break lines
% \usepackage{hyphenat}
\usepackage{bookmark}
\usepackage{systeme} % systems of equation
\usepackage[utf8]{inputenc} % necessary for proper unicode support
\usepackage{parcolumns} % multiple column parallel typesetting
\usepackage[english]{babel} % comment this for Angelsächsisch
\usepackage{graphicx} % including graphics
\usepackage{sourcecodepro} % best font
\usepackage{listings} % source code listings
\usepackage{biblatex} % better citation
% \usepackage[a4paper, top=1cm, left=2cm, right=2cm, includeheadfoot]{geometry} % customize page geometry

>>>>

\usepackage{amsmath, amsfonts, amssymb, amsthm}
\usepackage[normalem]{ulem} % underline (don't redefine emph to underline!)
\usepackage{float} % H float specifier
\usepackage{tabularx} % tabular with column specifier X to break lines
% \usepackage{hyphenat} % disable/enable hyphenation
\usepackage{bookmark} % this loads hyperref
\usepackage{multirow}
\usepackage{booktabs} % (top|mid|bottom)rule and other table stuff
\usepackage{fancyhdr} % fancy headers
\usepackage{systeme} % systems of equation
\usepackage[utf8]{inputenc} % necessary for proper unicode support
\usepackage{parcolumns} % multiple column parallel typesetting
\usepackage[english]{babel} % uncomment this for Angelsächsisch
\usepackage{graphicx} % including graphics
\usepackage{sourcecodepro} % best font
\usepackage{listings} % source code listings
\usepackage{caption} % customize captions in floats (rotate, sideways, ...)
\usepackage{lstlangarm} % custom ARM Assembler keyword list
\usepackage[backend=biber, sorting=none]{biblatex} % better citation
\usepackage[a4paper]{geometry} % customize page geometry

\pagestyle{fancy}
\fancyhead[L]{\rightmark}
\fancyhead[R]{\thepage}
\fancyfoot{}

% non-italic bibliography titles
\DeclareFieldFormat{title}{``#1''}
\DeclareFieldFormat{citetitle}{``#1''}
